
\chapter{Background}

Using computers to simulate computers is an obviously meta subject (perhaps to
the point of clich\'e). To avoid confusion it is common to make a distinction
between the \textit{target}, the computer being simulated, and the
\textit{host}, the computer executing (\textit{hosting}) the simulation. It is important
to note, the \textit{host} (or more illustrative, \textit{host-platform}) is
often not a single machine, but instead some collection of interconnected
machines, which may include CPUs, GPUs, and FPGAs, overwhich models of
different components of the target are hosted.


\section{A Brief Tour of Full-System Simulation}

One key driver of costs in the ASIC design process is simulation. Simulation preforms
three different functions.

\begin{enumerate}

    \item Prototyping: "What thing should we build?" Serves as a means to
        rapidly evaluate different design points with abstract models, or an
        incomplete implementation of a proposed design.

    \item Verification: "Did we build the thing right?" To check that a
        particular implementation has correctly executes.

    \item Validation: "Did we build the right thing?" To show that an
        implementation fulfills the high-level objectives set out for the
        system. Here the hope is that simulation of the implementation
        "validates" the results of the prototype.

\end{enumerate}

Radically different simulation technologies are deployed throughout the design
process. Each of these technologies makes a tradeoff between
\textit{fidelity}~(how accurate is the simulation), \textit{speed}~(how quickly
does simulator execute) and \textit{cost}~(\$ per simulation hour). That
operate, ranging from SPICE simulation of analog circuits, discrete event
simulation of RTL implementations, to architectural simulators that may do
litte or no accounting of time. \TODO{Some common technologies are illustrated
in Table X}

Few of these technologies are fast enough to host a simulation of the entire
design at speeds.

Generally, there are only two points during the development of an ASIC in which
\textit{full-system simulation}, executes within two orders of magnitude of.

\subsection{FPGA Accelerated Simulation}

\subsection{FPGA Accelerated Simulation is not FPGA Prototyping}

\section{An Overview of MIDAS}

\subsection{Host-Target Decoupling in MIDAS}

\subsection{Automatic FAME Transformations}

\subsection{Platform Mapping}

\subsection{Target Machines}

\section{An Overview of Off-Chip Memory Systems \& DRAM}

\subsection{DRAM Device Architecture}


\subsection{DRAM Memory Access Scheduling}


\subsection{System Integration of Off-Chip Memory Systems}
