\begin{tabularx}{\textwidth}{lX}
\multicolumn{2}{c}{\textbf{Abbreviations}} \\
\textbf{ASIC} & Application-Specific Integrated Circuit \\
\textbf{FAME} & FPGA Architecture Model Execution \\
\textbf{FPGA} & Field-Programmable Gate Array \\
\textbf{HDL} & Hardware Description Language \\
\textbf{LI-BDN} & Latency-Insensitive Bounded Dataflow Network \\
\textbf{LP} & Logical Process \\
\textbf{NRE} & Non-recurring Engineering [cost] \\
\textbf{PDES} & Parallel Discrete-Event Simulation \\
\textbf{SoC} & System-on-a-Chip. Used interchangeably with ASIC in this text. \\
\textbf{RTL} & Register Transfer Level \\
& \\
\multicolumn{2}{c}{\textbf{Key Definitions}} \\
\textbf{Bridge} & A target black-box and its custom unit implementation. \\
\textbf{Chisel} & An open-source HDL embedded in Scala \\
\textbf{Channel} & A point-to-point LP that may model stateful interconnect. \\
\textbf{FIRRTL} & Flexible Intermediate Representation for RTL \\
\textbf{Host Decoupling} & The property that permits simulated time to advance independently of wall-clock~(host) time.
Simulators are host-decoupled. FPGA prototypes are not. \\
\textbf{Host} & The system on which a simulation executes. \\
\textbf{Host Time} & Wall-clock time. Seen by the user and the host. \\
\textbf{Hub Unit} & The singleton unit representing all unoptimized RTL. \\
\textbf{Message} & A 2-tuple consisting of a signal value and a timestamp. \\
\textbf{Model} & An optimized UU. A valid primitive LI-BDN. \\
\textbf{Target} & The closed system under simulation. \\
\textbf{Target Time} & The time perceived by the system under simulation. See Target. \\
\textbf{Token} & An untimestamped data value. \\
\textbf{TU} & A timestamped unit; a unit that acts on messages. \\
%\textbf{FPGA Prototyping} & A fast, non-deterministic form of hardware-assisted verification in which
%a design is implemented directly on one or more FPGAs. \\
%\textbf{Hardware Emulation} & A more debuggable form of hardware-assisted verification in which
%target and host-time have been decoupled. Note, we use emulation and simulation interchangeably in this text. \\
\textbf{Unit} & An LP representing a part of an SoC or its environment. \\
\textbf{UU} & An untimestamped unit; a unit that acts on tokens. \\
\end{tabularx}
